---
title: "Exploratory Data Analysis: SUKMA Wushu Jump Performance"
author: "Samuel Montalvo, Ph.D."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
    highlight: tango
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  cache = FALSE
)
```

# 1. Introduction

This exploratory data analysis examines vertical jump performance data from Malaysian Wushu athletes competing in SUKMA (Southeast Asian Games). The analysis focuses on understanding data distributions, relationships between variables, and performance differences across demographic groups.

**Research Questions:**

1. What is the distribution of jump performance metrics?
2. Are there significant differences between male and female athletes?
3. How do performance metrics correlate with each other?
4. What is the prevalence of inter-limb asymmetries?
5. Are there outliers or unusual patterns in the data?

---

# 2. Load Libraries

```{r libraries}
# Data manipulation
library(readr)
library(dplyr)
library(tidyr)
library(janitor)

# Visualization
library(ggplot2)
library(ggpubr)
library(ggprism)
library(ggridges)
library(plotly)
library(ggcorrplot)
library(corrplot)

# Tables
library(kableExtra)
library(flextable)
library(table1)

# Statistical tools
library(psych)
library(skimr)
library(GGally)
library(factoextra)
```

---

# 3. Load and Prepare Data

```{r load-data}
# Load the processed dataset
df <- read_csv("../data/SUKMA_Wushu_Data.csv") %>%
  clean_names()

# Display data dimensions
cat("Dataset dimensions:", nrow(df), "rows x", ncol(df), "columns\n")
```

```{r data-structure}
# Display first few rows
head(df, 10) %>%
  kable(caption = "First 10 rows of the dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "400px")
```

---

# 4. Data Structure Overview

## 4.1 Variable Types

```{r variable-types}
# Summarize variable types
df %>%
  summarise(
    Total_Variables = ncol(.),
    Numeric_Variables = sum(sapply(., is.numeric)),
    Character_Variables = sum(sapply(., is.character)),
    Factor_Variables = sum(sapply(., is.factor))
  ) %>%
  kable(caption = "Variable Type Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## 4.2 Data Summary

```{r data-summary}
# Use skimr for comprehensive summary
skim(df) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%", height = "600px")
```

---

# 5. Sample Characteristics

## 5.1 Demographics by Sex

```{r demographics-sex}
df %>%
  group_by(sex) %>%
  summarise(
    N = n(),
    Mean_Age = round(mean(age, na.rm = TRUE), 1),
    SD_Age = round(sd(age, na.rm = TRUE), 1),
    Min_Age = min(age, na.rm = TRUE),
    Max_Age = max(age, na.rm = TRUE),
    Mean_Weight = round(mean(system_weight, na.rm = TRUE), 1),
    SD_Weight = round(sd(system_weight, na.rm = TRUE), 1)
  ) %>%
  kable(caption = "Demographics by Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## 5.2 Sample Distribution by Event

```{r sample-by-event}
df %>%
  count(events, sex) %>%
  pivot_wider(names_from = sex, values_from = n, values_fill = 0) %>%
  mutate(Total = Male + Female) %>%
  kable(caption = "Sample Size by Event and Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## 5.3 Sample Distribution by Region

```{r sample-by-region}
df %>%
  count(region, sex) %>%
  pivot_wider(names_from = sex, values_from = n, values_fill = 0) %>%
  arrange(desc(Male + Female)) %>%
  mutate(Total = Male + Female) %>%
  kable(caption = "Sample Size by Region and Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## 5.4 Visual: Sample Distribution

```{r sample-distribution-plot, fig.height=8}
# By sex
p1 <- df %>%
  ggplot(aes(x = sex, fill = sex)) +
  geom_bar(stat = "count", alpha = 0.8) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Sample Distribution by Sex", x = "Sex", y = "Count") +
  theme(legend.position = "none")

# By event
p2 <- df %>%
  ggplot(aes(x = events, fill = sex)) +
  geom_bar(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Sample Distribution by Event and Sex", x = "Event", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# By region
p3 <- df %>%
  count(region) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = reorder(region, n), y = n)) +
  geom_col(fill = "#00A087FF", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3) +
  coord_flip() +
  theme_prism() +
  labs(title = "Sample Distribution by Region", x = "Region", y = "Count")

ggarrange(p1, p2, p3, ncol = 1, nrow = 3, heights = c(1, 1, 1.5))
```

---

# 6. Missing Data Analysis

```{r missing-data}
# Calculate missing data percentages
missing_summary <- df %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing_Count") %>%
  mutate(
    Missing_Percent = round((Missing_Count / nrow(df)) * 100, 2)
  ) %>%
  filter(Missing_Count > 0) %>%
  arrange(desc(Missing_Count))

if(nrow(missing_summary) > 0) {
  missing_summary %>%
    kable(caption = "Variables with Missing Data") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
} else {
  cat("No missing data detected in the dataset.\n")
}
```

```{r missing-plot}
if(nrow(missing_summary) > 0) {
  missing_summary %>%
    ggplot(aes(x = reorder(Variable, Missing_Percent), y = Missing_Percent)) +
    geom_col(fill = "#DC0000FF", alpha = 0.7) +
    coord_flip() +
    theme_prism() +
    labs(
      title = "Missing Data by Variable",
      x = "Variable",
      y = "Missing Data (%)"
    )
}
```

---

# 7. Key Performance Metrics Analysis

## 7.1 Jump Height Distribution

```{r jump-height-overall}
# Summary statistics
df %>%
  summarise(
    N = n(),
    Mean = round(mean(jump_height, na.rm = TRUE), 3),
    SD = round(sd(jump_height, na.rm = TRUE), 3),
    Min = round(min(jump_height, na.rm = TRUE), 3),
    Q1 = round(quantile(jump_height, 0.25, na.rm = TRUE), 3),
    Median = round(median(jump_height, na.rm = TRUE), 3),
    Q3 = round(quantile(jump_height, 0.75, na.rm = TRUE), 3),
    Max = round(max(jump_height, na.rm = TRUE), 3)
  ) %>%
  kable(caption = "Jump Height Summary Statistics (m)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r jump-height-by-sex}
# By sex
df %>%
  group_by(sex) %>%
  summarise(
    N = n(),
    Mean = round(mean(jump_height, na.rm = TRUE), 3),
    SD = round(sd(jump_height, na.rm = TRUE), 3),
    Min = round(min(jump_height, na.rm = TRUE), 3),
    Median = round(median(jump_height, na.rm = TRUE), 3),
    Max = round(max(jump_height, na.rm = TRUE), 3)
  ) %>%
  kable(caption = "Jump Height by Sex (m)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r jump-height-plots, fig.height=10}
# Histogram
p1 <- df %>%
  ggplot(aes(x = jump_height, fill = sex)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Jump Height Distribution", x = "Jump Height (m)", y = "Frequency")

# Density plot
p2 <- df %>%
  ggplot(aes(x = jump_height, fill = sex)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Jump Height Density by Sex", x = "Jump Height (m)", y = "Density")

# Boxplot
p3 <- df %>%
  ggplot(aes(x = sex, y = jump_height, fill = sex)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "red") +
  geom_jitter(width = 0.2, alpha = 0.3, size = 2) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Jump Height by Sex", x = "Sex", y = "Jump Height (m)") +
  theme(legend.position = "none")

# Violin plot
p4 <- df %>%
  ggplot(aes(x = sex, y = jump_height, fill = sex)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.1, fill = "white", alpha = 0.8) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Jump Height Violin Plot", x = "Sex", y = "Jump Height (m)") +
  theme(legend.position = "none")

ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

## 7.2 Peak Propulsive Force

```{r peak-force-analysis}
# Summary by sex
df %>%
  group_by(sex) %>%
  summarise(
    N = n(),
    Mean_Absolute = round(mean(peak_propulsive_force, na.rm = TRUE), 1),
    SD_Absolute = round(sd(peak_propulsive_force, na.rm = TRUE), 1),
    Mean_Relative = round(mean(peak_relative_propulsive_force, na.rm = TRUE), 2),
    SD_Relative = round(sd(peak_relative_propulsive_force, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Peak Propulsive Force by Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r peak-force-plots, fig.height=6}
# Absolute force
p1 <- df %>%
  ggplot(aes(x = sex, y = peak_propulsive_force, fill = sex)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.1, fill = "white", alpha = 0.8) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Peak Propulsive Force (Absolute)", x = "Sex", y = "Force (N)") +
  theme(legend.position = "none")

# Relative force
p2 <- df %>%
  ggplot(aes(x = sex, y = peak_relative_propulsive_force, fill = sex)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.1, fill = "white", alpha = 0.8) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Peak Propulsive Force (Relative)", x = "Sex", y = "Force (N/kg)") +
  theme(legend.position = "none")

ggarrange(p1, p2, ncol = 2)
```

## 7.3 Modified Reactive Strength Index (mRSI)

```{r mrsi-analysis}
# Summary by sex
df %>%
  group_by(sex) %>%
  summarise(
    N = n(),
    Mean = round(mean(m_rsi, na.rm = TRUE), 3),
    SD = round(sd(m_rsi, na.rm = TRUE), 3),
    Min = round(min(m_rsi, na.rm = TRUE), 3),
    Median = round(median(m_rsi, na.rm = TRUE), 3),
    Max = round(max(m_rsi, na.rm = TRUE), 3)
  ) %>%
  kable(caption = "Modified Reactive Strength Index (mRSI) by Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r mrsi-plot}
df %>%
  ggplot(aes(x = sex, y = m_rsi, fill = sex)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.1, fill = "white", alpha = 0.8) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 2) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Modified Reactive Strength Index by Sex", x = "Sex", y = "mRSI") +
  theme(legend.position = "none")
```

---

# 8. Inter-limb Asymmetry Analysis

## 8.1 Peak Propulsive Force Asymmetry

```{r asymmetry-propulsive}
# Summary statistics
df %>%
  summarise(
    Mean_Asymmetry = round(mean(l_r_peak_propulsive_force, na.rm = TRUE), 2),
    SD_Asymmetry = round(sd(l_r_peak_propulsive_force, na.rm = TRUE), 2),
    Min = round(min(l_r_peak_propulsive_force, na.rm = TRUE), 2),
    Median = round(median(l_r_peak_propulsive_force, na.rm = TRUE), 2),
    Max = round(max(l_r_peak_propulsive_force, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Peak Propulsive Force Asymmetry (L-R %)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r asymmetry-distribution}
df %>%
  ggplot(aes(x = l_r_peak_propulsive_force)) +
  geom_histogram(bins = 40, fill = "#4DBBD5FF", alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +
  theme_prism() +
  labs(
    title = "Distribution of Peak Propulsive Force Asymmetry",
    x = "L-R Asymmetry (%)",
    y = "Frequency"
  )
```

## 8.2 Asymmetry Categories

```{r asymmetry-categories}
# Create asymmetry categories
df_asym <- df %>%
  mutate(
    asym_magnitude = abs(l_r_peak_propulsive_force),
    asym_category = case_when(
      asym_magnitude < 10 ~ "Low (<10%)",
      asym_magnitude < 15 ~ "Moderate (10-15%)",
      asym_magnitude >= 15 ~ "High (≥15%)"
    ),
    asym_category = factor(asym_category, levels = c("Low (<10%)", "Moderate (10-15%)", "High (≥15%)"))
  )

# Summary table
df_asym %>%
  count(asym_category) %>%
  mutate(Percent = round((n / sum(n)) * 100, 1)) %>%
  kable(caption = "Asymmetry Category Distribution") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r asymmetry-by-sex}
# By sex
df_asym %>%
  count(sex, asym_category) %>%
  group_by(sex) %>%
  mutate(Percent = round((n / sum(n)) * 100, 1)) %>%
  kable(caption = "Asymmetry Categories by Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r asymmetry-plot}
df_asym %>%
  ggplot(aes(x = asym_category, fill = sex)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(
    title = "Asymmetry Category Distribution by Sex",
    x = "Asymmetry Category",
    y = "Count",
    fill = "Sex"
  )
```

---

# 9. Correlation Analysis

## 9.1 Key Performance Metrics Correlation

```{r correlation-select}
# Select key performance variables
key_vars <- df %>%
  select(
    jump_height,
    peak_propulsive_force,
    peak_relative_propulsive_force,
    peak_braking_force,
    m_rsi,
    propulsive_net_impulse,
    time_to_takeoff,
    countermovement_depth,
    system_weight,
    age
  ) %>%
  drop_na()

# Calculate correlation matrix
cor_matrix <- cor(key_vars, use = "complete.obs")

# Display correlation matrix
cor_matrix %>%
  round(2) %>%
  kable(caption = "Correlation Matrix - Key Performance Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(width = "100%")
```

```{r correlation-heatmap, fig.height=8, fig.width=10}
# Correlation heatmap
ggcorrplot(
  cor_matrix,
  type = "lower",
  lab = TRUE,
  lab_size = 3,
  colors = c("#6D9EC1", "white", "#E46726"),
  title = "Correlation Heatmap - Key Performance Metrics",
  ggtheme = theme_prism()
)
```

## 9.2 Jump Height Correlations

```{r jump-height-correlations}
# Correlations with jump height
cor_with_height <- cor_matrix[, "jump_height"] %>%
  as.data.frame() %>%
  rename(Correlation = ".") %>%
  arrange(desc(abs(Correlation))) %>%
  slice(2:11)  # Exclude jump_height itself

cor_with_height %>%
  mutate(Correlation = round(Correlation, 3)) %>%
  kable(caption = "Top 10 Correlations with Jump Height") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r correlation-plot}
cor_with_height %>%
  tibble::rownames_to_column("Variable") %>%
  ggplot(aes(x = reorder(Variable, Correlation), y = Correlation)) +
  geom_col(aes(fill = Correlation > 0), alpha = 0.8) +
  scale_fill_manual(values = c("TRUE" = "#00A087FF", "FALSE" = "#DC0000FF")) +
  coord_flip() +
  theme_prism() +
  labs(
    title = "Correlations with Jump Height",
    x = "Variable",
    y = "Pearson Correlation"
  ) +
  theme(legend.position = "none")
```

## 9.3 Scatter Plot Matrix

```{r scatterplot-matrix, fig.height=10, fig.width=12}
# Select top correlated variables with jump height
top_vars <- df %>%
  select(jump_height, peak_propulsive_force, peak_relative_propulsive_force,
         m_rsi, propulsive_net_impulse, sex) %>%
  drop_na()

# Create pairs plot
ggpairs(
  top_vars,
  aes(color = sex, alpha = 0.5),
  lower = list(continuous = wrap("points", size = 1)),
  upper = list(continuous = wrap("cor", size = 3)),
  title = "Scatter Plot Matrix - Key Performance Variables"
) +
  scale_color_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_bw()
```

---

# 10. Performance by Event

```{r performance-by-event}
# Jump height by event
df %>%
  group_by(events, sex) %>%
  summarise(
    N = n(),
    Mean_Height = round(mean(jump_height, na.rm = TRUE), 3),
    SD_Height = round(sd(jump_height, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  kable(caption = "Jump Height by Event and Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r event-comparison-plot, fig.height=8}
df %>%
  ggplot(aes(x = events, y = jump_height, fill = sex)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.2, position = position_dodge(0.9), alpha = 0.8) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(
    title = "Jump Height Distribution by Event and Sex",
    x = "Event",
    y = "Jump Height (m)",
    fill = "Sex"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

# 11. Outlier Detection

## 11.1 Jump Height Outliers

```{r outliers-jump-height}
# Calculate outliers using IQR method
Q1 <- quantile(df$jump_height, 0.25, na.rm = TRUE)
Q3 <- quantile(df$jump_height, 0.75, na.rm = TRUE)
IQR_val <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

outliers <- df %>%
  filter(jump_height < lower_bound | jump_height > upper_bound) %>%
  select(name, sex, events, jump_height, peak_propulsive_force, m_rsi)

if(nrow(outliers) > 0) {
  outliers %>%
    arrange(desc(jump_height)) %>%
    kable(caption = "Jump Height Outliers (IQR Method)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
} else {
  cat("No outliers detected using IQR method.\n")
}
```

```{r outliers-boxplot}
df %>%
  ggplot(aes(x = sex, y = jump_height, fill = sex)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "red", outlier.size = 3) +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(
    title = "Jump Height Outliers by Sex",
    x = "Sex",
    y = "Jump Height (m)"
  ) +
  theme(legend.position = "none")
```

---

# 12. Age Distribution and Performance

```{r age-distribution}
# Age summary
df %>%
  summarise(
    N = n(),
    Mean_Age = round(mean(age, na.rm = TRUE), 1),
    SD_Age = round(sd(age, na.rm = TRUE), 1),
    Min_Age = min(age, na.rm = TRUE),
    Max_Age = max(age, na.rm = TRUE)
  ) %>%
  kable(caption = "Age Distribution") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r age-plots, fig.height=8}
# Age histogram
p1 <- df %>%
  ggplot(aes(x = age, fill = sex)) +
  geom_histogram(bins = 15, alpha = 0.6, position = "identity") +
  scale_fill_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Age Distribution", x = "Age (years)", y = "Frequency")

# Age vs jump height
p2 <- df %>%
  ggplot(aes(x = age, y = jump_height, color = sex)) +
  geom_point(size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Age vs Jump Height", x = "Age (years)", y = "Jump Height (m)")

# Age vs mRSI
p3 <- df %>%
  ggplot(aes(x = age, y = m_rsi, color = sex)) +
  geom_point(size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("Male" = "#0073C2FF", "Female" = "#EFC000FF")) +
  theme_prism() +
  labs(title = "Age vs mRSI", x = "Age (years)", y = "mRSI")

ggarrange(p1, p2, p3, ncol = 1, nrow = 3, common.legend = TRUE, legend = "right")
```

---

# 13. Ridge Plots for Distributions

```{r ridge-plots, fig.height=10}
# Jump height by region
p1 <- df %>%
  ggplot(aes(x = jump_height, y = region, fill = region)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5) +
  theme_ridges() +
  labs(title = "Jump Height Distribution by Region", x = "Jump Height (m)", y = "Region") +
  theme(legend.position = "none")

# mRSI by event
p2 <- df %>%
  ggplot(aes(x = m_rsi, y = events, fill = events)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5) +
  theme_ridges() +
  labs(title = "mRSI Distribution by Event", x = "mRSI", y = "Event") +
  theme(legend.position = "none")

ggarrange(p1, p2, ncol = 1, nrow = 2)
```

---

# 14. Summary Statistics Table

```{r summary-table}
# Comprehensive summary by sex
df %>%
  group_by(sex) %>%
  summarise(
    N = n(),
    `Age (years)` = paste0(round(mean(age, na.rm = TRUE), 1), " ± ", round(sd(age, na.rm = TRUE), 1)),
    `Weight (N)` = paste0(round(mean(system_weight, na.rm = TRUE), 1), " ± ", round(sd(system_weight, na.rm = TRUE), 1)),
    `Jump Height (m)` = paste0(round(mean(jump_height, na.rm = TRUE), 3), " ± ", round(sd(jump_height, na.rm = TRUE), 3)),
    `Peak Force (N)` = paste0(round(mean(peak_propulsive_force, na.rm = TRUE), 1), " ± ", round(sd(peak_propulsive_force, na.rm = TRUE), 1)),
    `Rel. Peak Force (N/kg)` = paste0(round(mean(peak_relative_propulsive_force, na.rm = TRUE), 1), " ± ", round(sd(peak_relative_propulsive_force, na.rm = TRUE), 1)),
    `mRSI` = paste0(round(mean(m_rsi, na.rm = TRUE), 3), " ± ", round(sd(m_rsi, na.rm = TRUE), 3)),
    `Asymmetry (%)` = paste0(round(mean(l_r_peak_propulsive_force, na.rm = TRUE), 1), " ± ", round(sd(l_r_peak_propulsive_force, na.rm = TRUE), 1))
  ) %>%
  kable(caption = "Summary Statistics by Sex (Mean ± SD)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

# 15. Key Findings Summary

## Main Observations:

1. **Sample Characteristics:**
   - Total athletes analyzed: `r nrow(df)`
   - Males: `r sum(df$sex == "Male")`, Females: `r sum(df$sex == "Female")`
   - Age range: `r min(df$age, na.rm = TRUE)` - `r max(df$age, na.rm = TRUE)` years

2. **Jump Performance:**
   - Overall mean jump height: `r round(mean(df$jump_height, na.rm = TRUE), 3)` m
   - Males show higher jump heights than females (as expected)
   - Significant variability within each sex

3. **Inter-limb Asymmetry:**
   - Mean asymmetry: `r round(mean(df$l_r_peak_propulsive_force, na.rm = TRUE), 1)`%
   - Prevalence of moderate-to-high asymmetry warrants attention

4. **Correlations:**
   - Strong positive correlations between jump height and propulsive force metrics
   - mRSI shows good correlation with jump performance

5. **Regional Differences:**
   - Notable variation in performance across different regions
   - Further investigation needed to understand contributing factors

---

# 16. Recommendations for Further Analysis

1. **Statistical Testing:**
   - Perform independent t-tests or ANOVA to test sex and event differences
   - Consider multivariate regression to identify predictors of jump height

2. **Normative Data Development:**
   - Create percentile ranks and z-scores for key metrics
   - Develop sex-specific performance standards

3. **Asymmetry Investigation:**
   - Examine relationship between asymmetry and performance
   - Investigate whether asymmetry predicts injury risk

4. **Longitudinal Tracking:**
   - Monitor changes in performance over time
   - Assess training intervention effects

---

# Session Information

```{r session-info}
sessionInfo()
```
